---
title: "Intelligent timeseries predictor"
author: "Anton Ehret"
output: html_document
---
```{r setup, include=FALSE}
#' @Setup
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/datensets"))

    library(forecast)
    library(stats)
    library(datasets)
    library(readxl)
    library(dplyr)
    library(plyr)
    library(ARDL) #for ADL analysis
    library(vars) #for VAR analysis
    library(tseries) #for ADF test
    library(lmtest) #for bp and dw test
    library(car) #for vif
    library(rio) #for changing xls into csv
    library(readr) #Fast csv write
```
Orange = Frontend input
Purple = Backend only
<h2>1. Upload of data by user</h2>
The Dataset has to be uploaded and meet the conditions of the first col being the dat and the 2nd col being the variable of interest. 
```{r data prep, include=FALSE}
#' *Name of dataset*
#Can this potentially be retreived automatically from the name of the uploaded dataset?
set <- "ARSales(1).xlsx"
#What do we do if we have a certain sheet that we are interested in?
# sheet <- "Datensatz1"

#' @load
  data.xlsx <- read_excel(set
                          # ,sheet = sheet
                          )

#' change into @csv for further manipulation and sorting
  data <- write_csv(data.xlsx, path="data.csv")
  
#' @rename first column for sorting
  colnames(data)[1] <- "Date"
  data$Date <- as.Date(data$Date, "%Y%m%d", optional = F) #turning char into date

#' @sort for dates in case data is sorted differently
  arrange(data, Date)
```

<h2>Input into Mask</h2>
```{r Mask}
#' *Variable of interest*
var <- data$Sales

#' *Split in %*
split <- 80

#' *Frequency`*
freq <- 12

#' *Start (Year)*
year <- 1998

#' *forecast period*
forc <- 12
```
<h2>Calculations are run in the background</h2>
A best model is chosen based on the lowest RMSE obtained. The models are trained on a training dataset and the accuracy is assessed in a test dataset.
The models which are trained are:
<ol>
<li> centered moving average </li>
<li> HoltWinters with different seasonal components</li>
<li> Autoregressive integrated Moving Average (ARIMA)</li>
</ol>
```{r ts and splitting data, message=FALSE, warning=FALSE, include=FALSE}
#' @ts variable of interest
  df <- ts(var
         , start = c(year, 01)  #' we start in year @2020
         , frequency = freq
         )

#'@splitting
  insamp <- ts(var[1:(length(var)*(split/100))]
               , start = c(year, 01)  #' we start in year @2020
               , frequency = freq
               )
  outsamp <- var[(length(var)*(split/100)) : length(var)]
```
```{r automatic Modelselection, message=FALSE, warning=FALSE, include=FALSE}
#' @models
#' Setting up
  h <- length(outsamp) #for forecasting period
  l <- list() #for models
  f <- list() #for forecasts
  a <- list() #for accuracy
#' Running models
  #Models
  l[[1]] <- ma(insamp
               , order = freq
               , centre = TRUE
               )
  l[[2]] <- ma(insamp
               , order = freq*2
               , centre = TRUE
               )
  l[[3]] <- HoltWinters(insamp)
  l[[4]] <- HoltWinters(insamp
                        , seasonal = "multiplicative"
                        )
  l[[5]] <- auto.arima(insamp)
  #Forecast & accuracy
  for (i in seq(1:5)){
    f[[i]] <- forecast(l[[i]]
                       , h = h
                       )
    a[[i]] <- accuracy(f[[i]]
                       , outsamp
                       )
  }

#' @model assessment
  #setting up sequence to only get the test-dataset assessment
  test <- seq(2,length(a)*2,2)
  #putting all accuraciy() in a dataframe
  assessment <- ldply(a,data.frame)[test,]
  #Modelselection based on RMSE
  chosen.model <- which.min(assessment$RMSE)
  
  if (chosen.model == 1) {
      mod <- ma(df, order = freq, centre = TRUE)
      message <- "Smallest RMSE was obtained by using centered moving average taking averages of one season"
    } else if (chosen.model == 2) {
      mod <- ma(df, order = freq*2, centre = TRUE)
      message <- "Smallest RMSE was obtained by using centered moving average taking averages of two seasons" 
    } else if (chosen.model == 3) {
      mod <- HoltWinters(df)
      message <- "Smallest RMSE was obtained by using HoltWinters with additive seasonal component"
    } else if (chosen.model == 4) {
      mod <- HoltWinters(df, seasonal = "multiplicative")
      message <- "Smallest RMSE was obtained by using HoltWinters with multiplicative seasonal component"
    } else {
      mod <- Arima(df, model=l[[5]])
      message <- "Smallest RMSE was obtained by using auto.arima"
    }
  
#' @Forecasting
    #' Forecast
    forecast <- forecast(mod, h = forc)
```
<h2>Forecast</h2>
```{r results, echo=FALSE}
#' *Displays in the frontend*
  #Message about which model was chosen
  print(message)
  #Forcasted plot
  plot(forecast, 
       main = "Forecast based on automatic selection of best Model")
   abline(v = year + length(df)/freq,
    col = "red")
    grid(col = "grey")
  forecast$model
```